Empowering the use of preprints in ecology and evolution: an exploration of the first 1000 *EcoEvoRxiv* preprints

### 2023-10-12 ms/ms.qmd
Updated MS with new paper

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used well. Significant barriers to preprint use still exist globally – from lack of clarity around preprint policies at journals [@Klebel2020] to a stigma that preprints are viewed by a research community to be of poor quality [@Chiarelli2019]. Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and allow us to begin to break down barriers to their use in our field.

Here, we explore the first 1000 preprints uploaded to *EcoEvoRxiv* to understand preprint practices and use in our field, focusing on our attention on understanding: 1) how long it takes for a preprint to become published, and how many preprints remain unpublished; 2) what countries make use of preprints; 3) how career stage and gender impact preprint use; 4) whether data and code are more likely to be shared in preprints; and 5) the extent to which authors make use of preprint servers for registered reports and community driven peer review.

>>>>

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used well. Significant barriers to preprint use still exist globally – from lack of clarity around preprint policies at journals [@Klebel2020] to a stigma that preprints are viewed by a research community to be of poor quality [@Chiarelli2019] (but see [@Faser2022]). Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and allow us to begin to break down barriers to their use in our field.

Here, we explore the first 1000 preprints uploaded to *EcoEvoRxiv* to understand preprint practices and use in our field, focusing our attention on understanding: 1) how long it takes for a preprint to become published, and how many preprints remain unpublished; 2) what countries make use of preprints; 3) how career stage and gender impact preprint use; 4) whether data and code are more likely to be shared in preprints; and 5) the extent to which authors make use of preprint servers for registered reports and community driven peer review.
